Copulas and their potential for ecology

### 2019-12-19 Paper.Rmd
finished trimming the main text

<!--Second mechanism, the Moran one-->
The hypothesis that characteristics of 
the copula structure of spatial dependence 
in an ecological variable can be inherited from an environmental 
variable was found to be reasonable, because it held for our models -
the Moran effect seems to extend to copula structure and tail associations - 
however, similarities between environmental- and 
ecological-variable copula structure were reduced when 
dynamics were strongly nonlinear.
For simulations using AR(1) models, our correlation and tail asymmetry statistics 
were always similar for noise and populations 
(Fig. \ref{fig_Cause4copula_C_kend_ar1} for the Clayton copula, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_ar1} for the survival Clayton copula).
Though there were significant differences for many statistics
and simulations, these were small compared to the overall tendency for
larger (respectively, smaller) values of our statistics, as applied to
noise, to be paired with larger (respectively, smaller) values of the same
statistics for model outputs. 
For our nonlinear model with weak noise, values of the statistics
were again quite similar for noise and model outputs 
(Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1} for Clayton, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_0.1} for survival Clayton).
Though there were again significant differences for many statistics
and simulations, these were again small relative to overall variation of 
values of the statistics.
Since many ecological models are nonlinear, this result provides the 
reasonable expectation for a Moran-effect-like correspondence 
between noise and model-output dependence structure across space 
for typical ecological dynamics,
as long as environmental noise is small enough that dynamics stay relatively close
to the model equilibrium. Theoretical results that hold for "weak noise" in this sense
are common in ecology.
For strong noise and using our nonlinear model, 
our correlation and tail-association-asymmetry statistics, generally 
speaking, were approximately similar between noise and model
outputs; however, similarity was reduced relative to previous results, 
and for a few simulations, asymmetry statistics
had opposite signs for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1} and \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_1}). For instance, 
using a Clayton  copula with a large Kendall $\tau$,
$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
population model outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}).

We repeated our analyses using the nonlinear model with $r=1.3$. The deterministic 
one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
at $K$ when $r<1$ (undercompensating dynamics, e.g., the value $r=0.5$ used initially), 
but exhibits an oscillatory approach to $K$ when $r>1$ (overcompensating
dynamics, e.g., $r=1.3$). For weak noise, similarities were again dominant between
values of our statistics on noise and population time series. 
For strong noise, however, discrepancies were often glaring. Apparently 
noise of standard deviation 1 interacted especially strongly with 
model nonlinearities when the model was in its overcompensatory 
regime. We repeated all analyses using the Gumbel, survival Gumbel, Joe, and 
survival Joe copulas, with substantially similar main conclusions (not shown).

@@

For our evolutionary model, the hypothesis was 
correct, for the models we employed, that asymmetric tail association in evolutionary changes 
can produce similarly asymmetric tail association between characters across
phylogeny tips.
Once character evolution was simulated 100 times for each of the
dependence structures we considered for evolutionary changes, 
we had 817 bivariate characters for each of 500 simulations (see section 
\ref{Causes}). We 
computed our nonparametric asymmetry statistics 
$\cor_{0,0.2}-\cor_{0.8,1}$, $\Ps_{0,0.2}-\Ps_{0.8,1}$,
and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ for each simulation output and 
produced a histogram for each statistic and for each
dependence structure (Fig. \ref{fig_cop_evo}). Results showed that 
asymmetries of tail associations in evolutionary changes
were associated with similar asymmetries of tail associations in extant characters.
Thus, based on our simulations, we cannot reject the possibility that
this is a proximate mechanism behind observed asymmetric tail associations
in our bird and mammal data.
This result is represented in Fig. \ref{MasterFigure} as the dashed box around 
"If character evolution has copula structure" and the arrow labelled "D".
This topic is revisited in the Discussion, as is the box "Asymmetric species
interactions" and the arrow labelled "C" in Fig. \ref{MasterFigure}.

@@

Recall that Q3 is: what are the consequences of non-normal copula
structure and asymmetric tail associations for ecological understanding and
applications?
We addressed Q3 by exploring, using both data and models, three hypothesized
ecological consequences of non-normal copula structure and asymmetric tail
associations. As for Q2, the models we used are intended as initial explorations, only,
of whether copula structure may have a meaningful influence on important ecological
phenomena. Therefore only simple models are used, and comprehensive explorations of the 
sensitivity of results to model strucuture and parameters are left for future work.

The hypothesis was presented in the Introduction that
the distribution (through time) of a spatially 
averaged quantity should be influenced by 
dependencies between the local quantities being averaged, including their
copula structure and tail associations. 
To make this hypothesis more precise, suppose an ecological variable 
$E_i(t)$ is measured at locations $i=1,\ldots,N$ and times $t=1,\ldots,T$, and
the spatial mean $\sum_i E_i(t)/N$ is of
interest. The $E_i(t)$ could be, for instance, local abundances of a pest or 
exploited species, or local fluxes of a greenhouse gas. If  $E_i$ and $E_j$ are 
associated primarily in their right tails for most location pairs $i$ and $j$, then
exceptionally large values tend to occur at the same time in most locations.
We hypothesize that this 
will tend to increase the skewness of the distribution 
of the spatial mean. Similarly, left-tail associations between local variables
should tend to decrease skewness. Strong positive skewness of the 
spatial-mean time series corresponds to "spikiness" of that time series,
i.e., occassional very large values, 
which corresponds to instability through time. 
The spatial-mean time series and its skewness may be quantities of 
principal importance for 
pest or resource abundance, for which extreme values (spikes) 
in the spatial mean may have large effects.

@@

To further address Q3, we also examined a hypothesis that 
asymmetric tail associations across space of an environmental variable can influence the 
extinction risk of a metapopulation. We hypothesized that environmental noises
exhibiting greater left-tail associations across different habitat patches would 
cause higher metapopulation extinction risks because then very bad years for the 
component populations would occur simultaneously in many patches,
reducing rescue effects. Here we assume, for simplicity, 
that low values of the environmental
variable are "bad" for the populations and high values are "good".
We tested the reasonableness of this hypothesis using a metapopulation
extension of the Lewontin-Cohen model,
$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
where the $i^{th}$ component of the length-$N$ vector 
$\vec{P}(t)$ represents
population density in the $i^{th}$ habitat patch at time $t$. 
The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
$\exp(r+\epsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
$\epsilon_i(t)$ represent environmental noises. They were standard-normally 
distributed, were independent through time, and showed the same 
spatial correlations for every simulation, 
but were made to exhibit stronger right- or left-tail associations between patches in
different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
matrix $D$ was a dispersal matrix
modelling local or global dispersal at rate $d$, in different simulations
(Appendix \ref{SM-ext_risk_model}).
After each step, 
if the density in a patch was $<1$, it was
set to $0$. We simulated the model 10000 times for each combination of parameters,
starting from $p_0 = 50$ in each patch, 
and calculated extinction risk after $25$ time steps.

Finally, and pursuing ideas that sprang from earlier theoretical work [@Reuman2017], we 
tested whether the copula structure of the dependences between population variables 
measured in different locations has consequences for the spatial version of Taylor's
law. Taylor's law is a commonly observed and widely studied 
[@Taylor1961; @Taylor1988; @Cohen2013; @Xu2016; @Reuman2017] 
empirical pattern that relates the 
variances of groups, $g$, of population measurements to the means of the 
groups via a power law, $v_g = a \times m_g^b$, or equivalently, 
$\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
the Taylor's law *exponent* or 
*slope*, and $\log(a)$ is the Taylor's law intercept. 
There are several versions of Taylor's law. For spatial Taylor's law,
given population density or abundance data $x_i(t)$ measured in locations
$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
So means $m_g$ and variances $v_g$ are computed across space. We refer to the matrix
with $x_i(t)$ in the $t^{\text{th}}$ row and the $i^{\text{th}}$ 
column as a *population matrix*. 
We consider that 
Taylor's law holds true for a dataset if the 
$\log(v_g)$ versus $\log(m_g)$ scatter plot 
for the dataset shows a linear, homoskedastic pattern.
Taylor's law has been 
verified empirically for hundreds of taxa and has been applied in numerous fields
including fisheries management, estimation of species persistence times, and
agriculture [@Cohen2015; @Reuman2017], so the question of whether copula
structure of spatial dependence influences Taylor's law may have applied
significance.

To explore the influence of copula structure and tail associations on spatial Taylor's
law, we carried out a series of simulations that generated sets of population 
matrices that had 
distinct copula structure between locations but that 
were otherwise statistically similar. For $k=1,\ldots,1000$
and X representing a Clayton, normal, Frank, or a survival Clayton copula family,
we generated $N \times n$ population matrices $m^{(\text{X},k)}$ with the following properties:
1) for any given values of $j$ and $k$, the values 
$\{ m_{t,j}^{(\text{X},k)} : t=1,\ldots,N  \}$ were exactly the same, as unordered
sets, for all X, i.e., the same actual population values were used for location $j$ 
in simulation $k$, regardless of X;
2) for any given $k$ and $j_1$ and $j_2$ such that $j_1 \neq j_2$, the Spearman correlations (computed through time)
$\cor_t(m_{t,j_1}^{(\text{X},k)},m_{t,j_2}^{(\text{X},k)})$ were the same, to within sampling
variation, i.e., correlations through time of populations in two locations were the same,
up to sampling variation, regardless of X;
3) the copula for the dependence between $m_{t,j_1}^{(\text{X},k)}$ and $m_{t,j_2}^{(\text{X},k)}$ was  from the family X.
We used $N=50$ and $n=25$. Details of how these matrices were generated are in 
Appendices \ref{SM-Consequence_TL_simulation} and \ref{SM-sect:copsurrognd}.
Thus for each $k$, the poulation matrices $m^{(\text{X},k)}$
for X taking the values Clayton, normal, Frank, and survival Clayton
were statistically similar except for the copula structure between locations, which was
X. Thus comparing how the $log(v_g)$ versus $\log(m_g)$ relationship may
manifest differently for $m^{(\text{X},k)}$
for different values of X constitutes a test of the influence of copula
structure on Taylor's law. We selected X equal to the Clayton, normal, Frank, and survival 
Clayton copulas to explore a range of tail association patterns.

@@

<!--what we did and why is it important to study?-->
We showed that non-normal copula structure and asymmetric tail associations are 
common across multiple sub-disciplines
in ecology, although these facets of ecological data are only occasionally 
taken into account [@Valpine2014; @Anderson2018; @Popovic2018]. We hypothesized 
mechanisms that may cause non-normal copula structures and asymmetric tail associations;
we discuss below how commonly some of our
mechanisms may operate. 
We also hypothesized and demonstrated important consequences of 
non-normal copula structure and asymmetric tail associations for the field of ecology. 
For instance, the skewness
of a spatial-average time series is influenced by asymmetric tail associations between its 
constituent time series: predominantly right-tail-associated local time series 
can lead to "spiky" spatially averaged time series, with large outbreaks; and 
predominantly left-tail-associated local time series can lead to spatially
averaged time series showing accentuated "crashes".
Thus tail associations could have implications for pest and exploited species.
Extinction risk and Taylor's law can also be altered by tail association 
patterns across space. 
In our view, our results make it reasonable to suggest 
that a more comprehensive understanding 
of many ecological phenomena may be possible if a
complete copula characterization of associations between variables is employed. 
Copula approaches are
well developed, statistically [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017], 
and have been introduced in accessible formats [@Anderson2018; @Genest2007]. Furthermore, 
open-source computer implementations exist (e.g., the `copula` and 
`VineCopula` packages in the R programming language). Ecologists can apply these tools
immediately. We created several interrelated randomization procedures
(Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
that have built upon existing copula methods.

<!--tropical ecosystems-->
The approaches we have demonstrated should apply equally well to 
data from tropical or temperate ecosystems.
There is no reason to expect that non-normal copula structure and 
asymmetric tail associations should be special properties of datasets from temperate 
regions. The mechanisms we have proposed of non-normal copula structure and asymmetric tail 
associations seem equally likely to apply anywhere. 
For instance, the Moran effect, which underpins 
one of our proposed mechanisms (Fig. \ref{MasterFigure}, B), is a standard mechanism
that occurs whenever environmental variables influence populations. And Liebig's
law and nonlinear environmental influences on ecosystems, which underpin 
another one of our proposed mechanisms (Fig. \ref{MasterFigure}, A), are widely
demonstrated phenomena that apply in all biomes.

@@

<!--a para on how likely one is to get copula structure via the Liebig mechanism-->
Our first proposed causal mechanism (Fig. \ref{MasterFigure}, A) may well operate commonly,
for two reasons.
First, Liebig's law and the idea of limiting 
nutrients are dominant paradigms in ecology, and 
many studies have documented nonlinear or 
threshold influences of environmental
variables on ecological quantities. 
Second, fluctuations in environmental variables through time
are very commonly correlated across space. 
Because these factors, which are the essential ingredients of
the mechanism, are so common, it is reasonable to hypothesize that the mechanism
may operate commonly. It may be a dominant cause of 
asymmetric tail associations and non-normal copula struture of ecological 
dependencies across space.
We provide some further support for the mechanism 
in our discussion of green spruce aphids and
winter temperature below.

<!--para on how likely one is to get copula structure via the Moran mechanism-->
There are also reasons to hypothesize that 
our Moran mechanism (Fig. \ref{MasterFigure}, B) 
may operate commonly: Moran effects are common [@Liebhold2004; @Sheppard2016; 
@Defriez2017_ocean; @Defriez2017_land], and 
non-normal copula structures and asymmetric tail associations are often found in 
environmental variables. If intense meteorological events are 
also widespread, then
environmental variables associated with these events should take extreme
values simultaneously across large spatial areas, producing tail associations in measurements
made through time at different locations. Non-extreme values may instead
be associated with local phenomena, and therefore may be
less correlated across large areas.
@Serinaldi2008 examined the spatial dependence 
of rainfall in Central Italy. Gumbel or Student 2-copulas 
were candidates for modelling dependence, and 
neither of these is a normal copula.
A long-term study (1950–2014) in the Loess Plateau of China [@She2018]
showed that a Gumbel copula effectively modeled 
the spatial dependence of drought variables. The Gumbel copula demonstrates 
asymmetric tail associations.
Bivariate copula analysis was also used in forecasting the 
co-occurrence of extreme events (flood or drought) 
over the North Sikkim Himalayas using spatial datasets [@Goswami2018].

@@

Without reliable branch length estimates, it is difficult to interpret the significance
    of the magnitude of changes in traits across a tree.
Developing tests of association based on copula structure may possibly lead the way to more
    robust methods for studying associations when we lack defensible estimates of
    branch lengths. 
We note that our phylogenetic analyses here consisted merely of simulations to assess
    whether interesting copula structure in a simple evolutionary process could 
    leave a detectable signal on the trait data for extant species.
Substantial work remains to be done before we have a copula-based method for 
    analyzing data on a phylogenetic tree. 
But the nonparametric nature of approaches
    based solely on data ranks, as many of our copula methods are, may be a promising avenue for 
    avoiding inaccuracies that arise via the highly structured model assumptions implicit
    in the method of phylogenetically independent contrasts and related methods.

@@

<!--Discuss stuff here on tradeoff and life history that Bever told me about-->
Relationships between BMR and body mass relate
to a trade-off between mass-specific BMR (BMR per unit body mass) and body
mass itself. Copulas probably interrelate with 
life-history trade-off theory in additional ways beyond what we demonstrated.
For instance, it is well known that energy allocation to a life function, F (e.g., reproduction)
will reduce the energy that can be allocated to other functions, $G_1$, $G_2$, $G_3$ 
(e.g., growth, predation avoidance). This is the principle of 
allocation. But F can trade off against any or all of 
the $G_i$. Therefore, for large F, approaching absolute 
limitations, there may be a strong 
association between F and $G_1$, for instance. For small F, there may be
little association because resources not allocated to F can instead be allocated 
to any combination of the $G_i$. This constitutes 
asymmetric tail association between F and $G_1$. @Winemiller1992
described a three-way trade-off in fishes between age of reproductive
maturity, juvenile survivorship, and fecundity. The trade-off should, in theory,
produce a tight association between age of maturity and fecundity for 
fishes with low age of maturity, but little such association for later-maturing
fishes because those fishes may invest the resources not invested in maturing quickly into
either fecundity or juvenile survival.
These ideas suggest that copulas may be 
useful for studying multi-dimensional trade-offs. But applications will 
require careful attention to the possible consequences of biased sampling:
if the degree of completeness of a dataset is associated with one or
more of the characters, then statistical artefacts
can bias conclusions (Appendix \ref{SM-sect:missingdata}, 
Fig. \ref{SM-fig:missingdata}). Multi-dimensional copulas may also be the 
appropriate copula approach
to studying multi-dimensional trade-offs. We used only bivariate 
analyses in this study solely for simplicity.
But statistical theory on multi-dimensional copulas is also well developed
[@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017], 
and open-source computer implementations exist for multivariate as well as 
bivariate copula methods (e.g., the `VineCopula`
package for the R programming language). Such approaches may be a 
useful next step in life-history theory and other areas.

>>>>

<!--Second mechanism, the Moran one-->
The hypothesis that characteristics of 
the copula structure of spatial dependence 
in an ecological variable can be inherited from an environmental 
variable was found to be reasonable, because it held for our models -
the Moran effect seems to extend to copula structure and tail associations - 
however, similarities between environmental- and 
ecological-variable copula structure were reduced when 
dynamics were strongly nonlinear.
For simulations using AR(1) models, our correlation and tail asymmetry statistics 
were always similar for noise and populations 
(Fig. \ref{fig_Cause4copula_C_kend_ar1} for the Clayton copula, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_ar1} for the survival Clayton copula).
Though there were significant differences for many statistics
and simulations, these were small compared to the overall tendency for
larger (respectively, smaller) values of our statistics, as applied to
noise, to be paired with larger (respectively, smaller) values of the same
statistics for model outputs. 
For our nonlinear model with weak noise, values of the statistics
were again quite similar for noise and model outputs 
(Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_0.1} for Clayton, 
Fig. \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_0.1} for survival Clayton).
Though there were again significant differences for many statistics
and simulations, these were again small relative to overall variation of 
values of the statistics.
Since many ecological models are nonlinear, this result provides the 
reasonable expectation for a Moran-effect-like correspondence 
between noise and model-output dependence structure across space 
for typical ecological dynamics,
as long as environmental noise is small enough that dynamics stay relatively close
to the model equilibrium. Theoretical results that hold for "weak noise" in this sense
are common in ecology.
For strong noise and using our nonlinear model, 
our correlation and tail-association-asymmetry statistics, generally 
speaking, were approximately similar between noise and model
outputs; however, similarity was reduced, 
and for a few simulations, asymmetry statistics
had opposite signs for noise and model outputs 
(Figs \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1} and \ref{SM-fig_Cause4copula_SC_kend_sRicker_nsd_1}). For instance, 
using a Clayton copula with a large Kendall $\tau$,
$\cor_l - \cor_u$ was slightly positive for noise, but slightly negative for 
population outputs (Fig. \ref{SM-fig_Cause4copula_C_kend_sRicker_nsd_1}).

We repeated our analyses using the nonlinear model with $r=1.3$. The deterministic 
one-habitat-patch Ricker model exhibits a monotonic approach to a stable equilibrium
at $K$ when $r<1$ (undercompensating dynamics, e.g., the value $r=0.5$ used initially), 
but exhibits an oscillatory approach when $r>1$ (overcompensating
dynamics, e.g., $r=1.3$). For weak noise, similarities were again dominant between
values of our statistics on noise and population time series. 
For strong noise, however, discrepancies were often glaring. Apparently 
noise of standard deviation 1 interacted especially strongly with 
model nonlinearities when the model was in its overcompensatory 
regime. We repeated all analyses using the Gumbel, survival Gumbel, Joe, and 
survival Joe copulas, with substantially similar main conclusions (not shown).

@@

For our evolutionary model, the hypothesis was 
correct, for the models we employed, that asymmetric tail association in evolutionary changes 
can produce similarly asymmetric tail association between characters across
phylogeny tips.
Once character evolution was simulated 100 times for each of the
dependence structures we considered for evolutionary changes, 
we had 817 bivariate characters for each of 500 simulations (see section 
\ref{Causes}). We 
computed our nonparametric asymmetry statistics 
$\cor_{0,0.2}-\cor_{0.8,1}$, $\Ps_{0,0.2}-\Ps_{0.8,1}$,
and $\Dsq_{0.8,1}-\Dsq_{0,0.2}$ for each simulation output and 
produced a histogram for each statistic and for each
dependence structure (Fig. \ref{fig_cop_evo}). Results showed that 
asymmetries of tail associations in evolutionary changes
were associated with similar asymmetries of tail associations in extant characters.
Thus we cannot reject the possibility that
this is a proximate mechanism behind observed asymmetric tail associations
in our bird and mammal data.
This result is represented in Fig. \ref{MasterFigure} as the dashed box around 
"If character evolution has copula structure" and the arrow labelled "D".
This topic is revisited in the Discussion, as is the box "Asymmetric species
interactions" and the arrow labelled "C" in Fig. \ref{MasterFigure}.

@@

Recall that Q3 is: what are the consequences of non-normal copula
structure and asymmetric tail associations for ecological understanding and
applications?
We addressed Q3 by exploring, using both data and models, three hypothesized
consequences. As for Q2, the models we used are intended as initial explorations, only. 
Therefore only simple models are used, and comprehensive explorations of the 
sensitivity of results to model strucuture and parameters are left for future work.

The hypothesis was presented in the Introduction that
the distribution (through time) of a spatially 
averaged quantity should be influenced by 
dependencies between the local quantities being averaged, including their
copula structure and tail associations. 
To make this hypothesis more precise, suppose an ecological variable 
$E_i(t)$ is measured at locations $i=1,\ldots,N$ and times $t=1,\ldots,T$, and
the spatial mean $\sum_i E_i(t)/N$ is of
interest. The $E_i(t)$ could be, for instance, local abundances of a pest or 
exploited species, or local fluxes of a greenhouse gas. If  $E_i$ and $E_j$ are 
associated primarily in their right tails for most location pairs $i$ and $j$, then
exceptionally large values tend to occur at the same time in most locations.
We hypothesize that this can increase the skewness of the distribution 
of the spatial mean. Similarly, left-tail associations between local variables
should decrease skewness. Strong positive skewness of the 
spatial-mean time series corresponds to "spikiness" of that time series,
i.e., occassional very large values, 
which corresponds to instability through time. 
The spatial-mean time series and its skewness may be quantities of 
principal importance for 
pest or resource abundance, for which extreme values (spikes) 
in the spatial mean may have large effects.

@@

We also examined a hypothesis that 
asymmetric tail associations across space of an environmental variable can influence the 
extinction risk of a metapopulation. We hypothesized that environmental noises
exhibiting greater left-tail associations across habitat patches would 
cause higher metapopulation extinction risks because then very bad years for the 
component populations occur simultaneously in many patches,
reducing rescue effects. Here we assume, for simplicity, 
that low values of the environmental
variable are "bad" for the populations and high values are "good".
We tested the reasonableness of this hypothesis using a metapopulation
extension of the Lewontin-Cohen model,
$\vec{P}(t+1) = D \lambda(t) \vec{P}(t)$,
where the $i^{th}$ component of the length-$N$ vector 
$\vec{P}(t)$ represents
population density in the $i^{th}$ habitat patch at time $t$. 
The $N \times N$ matrix $\lambda(t)$ was diagonal with $i^{th}$ diagonal entry 
$\exp(r+\epsilon_i(t))$. Here $r$ is a growth rate; we used $r=0$. The
$\epsilon_i(t)$ represent environmental noises. They were standard-normally 
distributed, were independent through time, and showed the same 
spatial correlations for every simulation, 
but were made to exhibit stronger right- or left-tail associations between patches in
different simulations (Appendix \ref{SM-ext_risk_model}). The $N \times N$
matrix $D$ was a dispersal matrix
modelling local or global dispersal at rate $d$, in different simulations
(Appendix \ref{SM-ext_risk_model}).
After each step, 
if the density in a patch was $<1$, it was
set to $0$. We simulated the model 10000 times for each combination of parameters,
starting from $p_0 = 50$ in each patch, 
and calculated extinction risk after $25$ time steps.

Finally, and pursuing ideas from @Reuman2017, we 
tested whether the copula structure of the dependences between population variables 
measured in different locations has consequences for the spatial version of Taylor's
law. Taylor's law is a commonly observed and widely studied 
[@Taylor1961; @Taylor1988; @Cohen2013; @Xu2016; @Reuman2017] 
empirical pattern that relates the 
variances of groups, $g$, of population measurements to the means of the 
groups via a power law, $v_g = a \times m_g^b$, or equivalently, 
$\log(v_g)=\log(a)+b \times \log(m_g)$. Here $b$ is called 
the Taylor's law *exponent* or 
*slope*, and $\log(a)$ is the Taylor's law intercept. 
There are several versions of Taylor's law. For spatial Taylor's law,
given population density or abundance data $x_i(t)$ measured in locations
$i=1,\ldots,N$ and times $t=1,\ldots,T$, a group $g$ consists of 
all the measurements $x_i(t)$, $i=1,\ldots,N$ made in different locations at the same time.
So means $m_g$ and variances $v_g$ are computed across space. We refer to the matrix
with $x_i(t)$ in the $t^{\text{th}}$ row and the $i^{\text{th}}$ 
column as a *population matrix*. 
We consider that 
Taylor's law holds true for a dataset if the 
$\log(v_g)$ versus $\log(m_g)$ scatter plot 
for the dataset shows a linear, homoskedastic pattern.
Taylor's law has been 
verified empirically for hundreds of taxa and has been applied in numerous fields
including fisheries management, estimation of species persistence times, and
agriculture [@Cohen2015; @Reuman2017], so the question of whether copula
structure of spatial dependence influences Taylor's law may have applied
significance.

To explore the influence of copula structure and tail associations on spatial Taylor's
law, we carried out a series of simulations that generated sets of population 
matrices that had 
distinct copula structure between locations but that 
were otherwise statistically similar. For $k=1,\ldots,1000$
and X representing a Clayton, normal, Frank, or a survival Clayton copula family,
we generated $N \times n$ population matrices $m^{(\text{X},k)}$ with the following properties:
1) for any given values of $j$ and $k$, the values 
$\{ m_{t,j}^{(\text{X},k)} : t=1,\ldots,N  \}$ were exactly the same, as unordered
sets, for all X, i.e., the same actual population values were used for location $j$ 
in simulation $k$, regardless of X;
2) for any given $k$, and $j_1$ and $j_2$ such that $j_1 \neq j_2$, the 
Spearman correlations (computed through time)
$\cor_t(m_{t,j_1}^{(\text{X},k)},m_{t,j_2}^{(\text{X},k)})$ were the same, to within sampling
variation, i.e., correlations through time of populations in two locations were the same,
up to sampling variation, regardless of X;
3) the copula for the dependence between $m_{t,j_1}^{(\text{X},k)}$ and $m_{t,j_2}^{(\text{X},k)}$ was  from the family X.
We used $N=50$ and $n=25$. Details of how these matrices were generated are in 
Appendices \ref{SM-Consequence_TL_simulation} and \ref{SM-sect:copsurrognd}.
Thus for each $k$, the poulation matrices $m^{(\text{X},k)}$
for X taking the values Clayton, normal, Frank, and survival Clayton
were statistically similar except for the copula structure between locations, which was
X. Thus comparing how the $log(v_g)$ versus $\log(m_g)$ relationship may
manifest differently for $m^{(\text{X},k)}$
for different values of X constitutes a test of the influence of copula
structure on Taylor's law. We selected X equal to the Clayton, normal, Frank, and survival 
Clayton copulas to explore a range of tail association patterns.

@@

<!--what we did and why is it important to study?-->
We showed that non-normal copula structure and asymmetric tail associations are 
common across multiple sub-disciplines
in ecology, although these facets of data are only occasionally 
accounted for [@Valpine2014; @Anderson2018; @Popovic2018]. We hypothesized 
mechanisms that may cause non-normal copula structures and asymmetric tail associations;
we discuss below how commonly some of our
mechanisms may operate. 
We also demonstrated important consequences of 
non-normal copula structure and asymmetric tail associations for ecology. 
For instance, the skewness
of a spatial-average time series is influenced by asymmetric tail associations between its 
constituent time series: predominantly right-tail-associated local time series 
can lead to "spiky" spatially averaged time series, with large outbreaks; and 
predominantly left-tail-associated local time series can lead to spatially
averaged time series showing accentuated "crashes".
Thus tail associations could have implications for pests and exploited species.
Extinction risk and Taylor's law can also be altered by tail association 
patterns across space. 
In our view, our results make it reasonable to suggest 
that a more comprehensive understanding 
of many ecological phenomena may be possible if a
complete copula characterization of associations between variables is employed. 
Copula statistics are
well developed [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017], 
and have been introduced in accessible formats [@Anderson2018; @Genest2007]. Furthermore, 
open-source computer implementations exist (e.g., the `copula` and 
`VineCopula` packages in R). Ecologists can apply these tools
immediately. We created several interrelated randomization procedures
(Appendices \ref{SM-surrog_test}, \ref{SM-Surrog_multi}, \ref{SM-sect:copsurrognd}) 
that built upon existing copula methods.

<!--tropical ecosystems-->
The approaches we demonstrated should apply equally well to 
data from tropical or temperate ecosystems.
There is no reason to expect that non-normal copula structure and 
asymmetric tail associations should be special properties of datasets from temperate 
regions. The mechanisms we proposed of non-normal copula structure and asymmetric tail 
associations seem equally likely to apply anywhere. 
For instance, the Moran effect, which underpins 
one of our proposed mechanisms (Fig. \ref{MasterFigure}, B), is a standard mechanism
that occurs whenever environmental variables influence populations. And Liebig's
law and nonlinear environmental influences on ecosystems, which underpin 
another one of our mechanisms (Fig. \ref{MasterFigure}, A), are widely
demonstrated phenomena.

@@

<!--a para on how likely one is to get copula structure via the Liebig mechanism-->
Our first proposed causal mechanism (Fig. \ref{MasterFigure}, A) may well operate commonly,
for two reasons.
First, Liebig's law and the idea of limiting 
nutrients are dominant paradigms in ecology, and 
many studies have documented nonlinear or 
threshold influences of environmental
variables on ecological quantities. 
Second, fluctuations in environmental variables through time
are very commonly correlated across space. 
Because these factors, which are the essential ingredients of
the mechanism, are common, it is reasonable to hypothesize that the mechanism
may operate commonly and may be a dominant cause of 
asymmetric tail associations and non-normal copula struture of ecological 
dependencies across space.
We provide further support for the mechanism 
in our discussion of green spruce aphids and
winter temperature below.

<!--para on how likely one is to get copula structure via the Moran mechanism-->
There are also reasons to hypothesize that 
our Moran mechanism (Fig. \ref{MasterFigure}, B) 
may operate commonly: Moran effects are common [@Liebhold2004; @Sheppard2016; 
@Defriez2017_ocean; @Defriez2017_land], and 
non-normal copula structures and asymmetric tail associations are often found in 
environmental variables. If intense meteorological events are 
also widespread, then
environmental variables associated with these events should take extreme
values simultaneously across large spatial areas, producing tail associations in measurements
made through time at different locations. Non-extreme values may instead
be associated with local phenomena, and therefore may be
less correlated across large areas.
@Serinaldi2008 examined the spatial dependence 
of rainfall in Central Italy. Gumbel or Student 2-copulas 
were candidates for modelling dependence, and 
neither of these is a normal copula.
A long-term study (1950–2014) in the Loess Plateau of China [@She2018]
showed that a Gumbel copula effectively modeled 
the spatial dependence of drought variables. The Gumbel copula has 
asymmetric tail association.
Bivariate copula analysis was also used in forecasting the 
co-occurrence of extreme events (flood or drought) 
over the North Sikkim Himalayas using spatial datasets [@Goswami2018].

@@

Without reliable branch length estimates, it is difficult to interpret the significance
    of the magnitude of changes in traits across a tree.
Developing tests of association based on copula structure may possibly lead the way to more
    robust methods for studying associations when we lack defensible estimates of
    branch lengths. 
We note that our phylogenetic analyses here consisted merely of simulations to assess
    whether interesting copula structure in a simple evolutionary process could 
    leave a detectable signal on the trait data for extant species.
Substantial work remains to be done before we have a copula-based method for 
    analyzing data on a phylogenetic tree. 
But the nonparametric nature of approaches
    based solely on data ranks, as many of our methods are, may be a promising avenue for 
    avoiding inaccuracies that arise via the highly structured model assumptions implicit
    in the method of phylogenetically independent contrasts and related methods.

@@

<!--Discuss stuff here on tradeoff and life history that Bever told me about-->
Relationships between BMR and body mass relate
to a trade-off between mass-specific BMR (BMR per unit body mass) and body
mass itself. Copulas probably interrelate with 
life-history trade-off theory in additional ways beyond what we demonstrated.
For instance, it is well known that energy allocation to a life function, F (e.g., reproduction)
will reduce the energy that can be allocated to other functions, $G_1$, $G_2$, $G_3$ 
(e.g., growth, predation avoidance). This is the principle of 
allocation. But F can trade off against any or all of 
the $G_i$. Therefore, for large F, approaching absolute 
limitations, there may be a strong 
association between F and $G_1$, for instance. For small F, there may be
little association because resources not allocated to F can instead be allocated 
to any combination of the $G_i$. This constitutes 
asymmetric tail association between F and $G_1$. @Winemiller1992
described a three-way trade-off in fishes between age of reproductive
maturity, juvenile survivorship, and fecundity. The trade-off should, in theory,
produce a tight association between age of maturity and fecundity for 
fishes with low age of maturity, but little such association for later-maturing
fishes because those species may invest the resources not invested in maturing quickly into
either fecundity or juvenile survival.
These ideas suggest that copulas may be 
useful for studying multi-dimensional life-history trade-offs. But applications will 
require careful attention to the possible consequences of biased sampling:
if the degree of completeness of a dataset is associated with one or
more of the characters, then statistical artefacts
can bias conclusions (Appendix \ref{SM-sect:missingdata}, 
Fig. \ref{SM-fig:missingdata}). Multi-dimensional copulas may also be the 
appropriate copula approach
to studying multi-dimensional trade-offs. We used only bivariate 
analyses in this study solely for simplicity.
But statistical theory on multi-dimensional copulas is also well developed
[@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017], 
and open-source computer implementations exist for multivariate as well as 
bivariate copula methods (e.g., the `VineCopula`
package for R). Such approaches may be a 
useful next step in life-history theory.
